---
title: "Flight Price Prediction"
Author: Alfred Cheung

execute: 
  message: false
  echo: false
  warning: false
  
---


```{r}
library(tidyverse)
library(readxl)             
library(tidymodels)     
library(broom)          
library(marginaleffects) 
library(easystats)
library(stringr)
```

```{r}
x <- read_xlsx("Data_Train.xlsx")

# Use a R code to do the mutate

x1 <- x %>%
  filter(Airline != "Jet Airways Business")|>
  mutate(
    duration_hr = str_extract(Duration, "\\d+h") %>%             # 抓 "2h"
      str_remove("h") %>%                                        # 去除 h
      as.numeric(),
    
    duration_min = str_extract(Duration, "\\d+m") %>%            # 抓 "50m"
      str_remove("m") %>%                                        # 去除 m
      as.numeric(),
    
    duration_hr = if_else(is.na(duration_hr), 0, duration_hr),   # 沒有小時時補 0
    duration_min = if_else(is.na(duration_min), 0, duration_min),# 沒有分鐘時補 0
    
    duration_total_hr = duration_hr + duration_min / 60          # 合併為小時數
  )

```


## Average Flight Price by Airline


```{r}
library(scales)

ggplot(x1, aes(x = reorder(Airline, Price, FUN = mean), y = Price, fill = Airline)) +
  stat_summary(fun = mean, geom = "bar", alpha = 0.7) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  labs(
    x = "Airline",
    y = "Average Price"
  ) +
  scale_y_continuous(labels = dollar) +  # <- this adds $ and comma formatting
  theme_minimal(base_size = 14) +
  theme(legend.position = "none") +
  coord_flip()

```

The bar chart displays the average flight prices by airline, along with error bars indicating price variability. Jet Airways, Multiple carriers Premium economy, and Air India show the highest average prices among all airlines. In contrast, Trujet, SpiceJet, and Air Asia offer the lowest average fares. Notably, Vistara Premium economy has a high variability in pricing, suggesting inconsistency or limited sample size.



## Average Flight Price by Number of Stops

```{r}
# Assuming your data is in a dataframe called 'x1'
# If not, replace 'x1' with your actual dataframe name




x2 <- x1 %>%
  filter(Total_Stops!= c("4 stops", "NA"))

# Create a boxplot showing price distribution by total stops
p1 <- ggplot(x2, aes(x = Total_Stops, y = Price, fill = Total_Stops)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 0.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Flight Price Distribution by Number of Stops",
       subtitle = "Boxplot with individual data points",
       x = "Number of Stops",
       y = "Price (₹)",
       fill = "Stops") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  scale_fill_brewer(palette = "Set3")

# Create a violin plot for better distribution visualization
p2 <- ggplot(x2, aes(x = Total_Stops, y = Price, fill = Total_Stops)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.1, alpha = 0.8, outlier.shape = NA) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Flight Price Distribution by Number of Stops",
       subtitle = "Violin plot showing price density",
       x = "Number of Stops",
       y = "Price",
       fill = "Stops") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  scale_fill_brewer(palette = "Set2")

# Create summary statistics
price_summary <- x2 %>%
  group_by(Total_Stops) %>%
  summarise(
    count = n(),
    mean_price = mean(Price, na.rm = TRUE),
    median_price = median(Price, na.rm = TRUE),
    sd_price = sd(Price, na.rm = TRUE),
    min_price = min(Price, na.rm = TRUE),
    max_price = max(Price, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_price))

# Create a bar plot of average prices
p3 <- ggplot(price_summary, aes(x = reorder(Total_Stops, mean_price), y = mean_price, fill = Total_Stops)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = paste0(round(mean_price, 0))), 
            hjust = -0.1, size = 3) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Average Flight Price by Number of Stops",
       subtitle = "Mean prices with values labeled",
       x = "Number of Stops",
       y = "Average Price",
       fill = "Stops") +
  theme_minimal() +
  coord_flip() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set1")

# Display the plots
print(p2)
print(p3)
# Print summary statistics
cat("\nSummary Statistics by Total Stops:\n")
print(price_summary)

# Additional analysis - correlation if we convert stops to numeric
stops_numeric <- case_when(
  x1$Total_Stops == "non-stop" ~ 0,
  x1$Total_Stops == "1 stop" ~ 1,
  x1$Total_Stops == "2 stops" ~ 2,
  x1$Total_Stops == "3 stops" ~ 3,
  x1$Total_Stops == "4 stops" ~ 4,
  TRUE ~ NA_real_
)

correlation <- cor(stops_numeric, x1$Price, use = "complete.obs")
cat(paste("\nCorrelation between number of stops and price:", round(correlation, 3)))
```


Flight prices tend to increase with the number of stops. Non-stop flights are the cheapest on average, with a mean price of $5,025, while flights with 3 stops are the most expensive, averaging $13,112. The correlation coefficient of 0.621 indicates a moderately strong positive relationship between the number of stops and ticket price. This suggests that more stops generally lead to higher fares.
